---
title: "DK Nationwide LiDAR 'process_tiles.py' Quality Assurance Report"
author: "Jakob Assmann"
date: "30/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r prep, include = FALSE}
library(dplyr)
library(tidyverse)

# Load report data
overall_progress <- 
  read.csv("D:/Jakob/ecodes-dk-lidar-rev1/log/process_tiles/overall_progress.csv",
           stringsAsFactors = F) %>%
  filter(processing == "complete")

# Extract performance stats for each function
n_success <- sapply(
  colnames(overall_progress), 
  function(x){
    sum(overall_progress[,x] == "success")
  })
n_complete <- sapply(
  colnames(overall_progress), 
  function(x){
    sum(overall_progress[,x] == "complete")
  })
n_success["processing"] <- n_complete["processing"]
# Remove tile id and validate crs colums
n_success <- n_success[!(grepl("tile_id", names(n_success)) |
                            grepl("odm_validate_crs", names(n_success)) |
                           grepl("dtm_validate_crs", names(n_success)))]

# get tile_ids reports for tiles that were not succesffuly completed
steps_not_completed <- names(n_success[n_success != nrow(overall_progress)])


# Obtain error codes
errors <- lapply(
  steps_not_completed,
  function(x){
    print(x)
    print(unique(overall_progress[,x]))
  }
) %>% unlist() %>%
  unique()
errors <- errors[!(errors %in% c("success", steps_not_completed))]

# subset only rows with errors
tiles_with_errors <- apply(overall_progress, 1, function(x){
  if(sum(x %in% errors) >= 1) {
    cat(x[x %in% errors])
    return(x[1])
    }
  else return(0)
})
tiles_with_errors <- tiles_with_errors[tiles_with_errors != "0"]

incomplete_processing <- overall_progress[overall_progress$tile_id %in% tiles_with_errors,]

```

## Summary

A __total of `r nrow(overall_progress)`__ tiles were processed, of these __`r sum(overall_progress$processing == "complete")` tiles (`r round(sum(overall_progress$processing == "complete")/nrow(overall_progress)*100,2)`%) completed__  the processing.  

```{r processing success}
library(knitr)
incomplete_processing %>%
  filter(processing != "complete") %>%
  select(tile_id, processing) %>%
  kable(caption = "Tiles where processing failed completely:")
```

## Incomplete processing steps

For __`r nrow(incomplete_processing)` tiles (`r round(nrow(incomplete_processing)/nrow(overall_progress)*100,2)`%)__ one or multiple of the indiviudal processing steps could not be completed.

```{r incomplete_processing}
library(knitr)
incomplete_processing %>% 
  filter(processing == "complete") %>% 
  kable(caption = "Tiles with incomplete processing steps:")
```

## Projections

Of all tiles, __`r length(grep('empty - set', overall_progress$odm_validate_crs))` laz tiles (`r round(length(grep('empty - set', overall_progress$odm_validate_crs))/nrow(overall_progress)*100,2)`%)__ did not have a CRS set and __`r length(grep('empty - set', overall_progress$drm_validate_crs))` dtm tiles (`r round(length(grep('empty - set', overall_progress$dtm_validate_crs))/nrow(overall_progress)*100,2)`%)__ did not have a CRS set.

For __`r length(grep('no match', overall_progress$odm_validate_crs))` laz
tiles (`r round(length(grep('no match', overall_progress$odm_validate_crs))/nrow(overall_progress)*100,2)`%)__ the crs did not match the default crs and __`r length(grep('no match', overall_progress$drm_validate_crs))` dtm tiles (`r round(length(grep('no match', overall_progress$odm_validate_crs))/nrow(overall_progress)*100,2)`%)__ the crs
did not match the defualt crs.

```{r crs_not_matched}
try(overall_progress[grep('no match', overall_progress$dtm_validate_crs),] %>%
  select(tile_id, odm_validate_crs) %>%
    slice(1:100) %>%
  kable(caption = "Tiles with an laz file that has a non-default crs (first 100):"))

try(overall_progress[grep('no match', overall_progress$dtm_validate_crs),] %>%
  select(tile_id, dtm_validate_crs) %>%
    slice(1:100) %>%
  kable(caption = "Tiles with a dtm file that has a non-default crs (first 100):"))
```

